Lens Modeling in \sw

### 2013-09-10 text/tex/results.tex

@@

T1 was also designed to give some feedback on the difficulties volunteers encounter, to further improve the tutorial materials.

@@

\begin{figure}
  \centering
  \subfigure{\includegraphics[width=0.45\textwidth]
             {fig/sims/006917/input.png}}
  \subfigure{\includegraphics[width=0.45\textwidth]
             {fig/sims/006917/spaghetti.png}} \\
  \subfigure{\includegraphics[width=0.45\textwidth]
             {fig/sims/ASW0001hpf/arriv.png}}
  \subfigure{\includegraphics[width=0.45\textwidth]
             {fig/sims/ASW0001hpf/extr_points.png}}
  \caption{Examples of Spaghetti input (left top) and modeled arrival time surface (right top) vs reconstructed arrival time surface from simulation parameters (bottom left), additionally using a extremal point detector (bottom right).}
  \label{fig:output_compare}
\end{figure}

@@

\hr

\todo{incl this figs}
Figure \ref{fig:6941} shows a simple example.

>>>>

\subsection{Some example models}

Figure \ref{fig:6941} shows a simple example.

@@

T1 was also designed to give some feedback on the difficulties volunteers encounter, to further imrove the tutorial materials.

@@

\begin{figure}
  \centering
  \subfigure{\includegraphics[width=0.45\textwidth]
             {fig/sims/006917/input.png}}
  \subfigure{\includegraphics[width=0.45\textwidth]
             {fig/sims/006917/spaghetti.png}} \\
  \subfigure{\includegraphics[width=0.45\textwidth]
             {fig/sims/ASW0001hpf/arriv.png}}
%  \subfigure{\includegraphics[width=0.45\textwidth]
%             {fig/sims/ASW0001hpf/extr_points.png}}
  \caption{Examples of Spaghetti input (left top) and modeled arrival time surface (right top) vs reconstructed arrival time surface from simulation parameters (bottom left), additionally using a extremal point detector (bottom right).}
  \label{fig:output_compare}
\end{figure}

@@

### 2013-09-10 text/tex/spaghetti.tex

In \spl the user begins by making an educated guess for the topography
of the arrival-time surface --- specifically, the locations of the
maxima, minima and saddle points that correspond to images of the
brightest part of the source.  Input is given by tracing a proposal
for the saddle-point contours.  Figure \ref{fig:input-spag} shows some
example inputs --- and incidentally also indicates the origin of the
name \spl.

@@

\spl provides plug ins to support any data source. At the moment, a data link to \sw and to the MasterLens database\needcite are implemented.

\spl assists volunteers with several features in this process.
Step 1 by supplying several images from several bands (not yet implemented in \sw).

Give a guess for the maximum, minimum and saddle points.  Program
tries find $\kappa(x,y)$ that reproduces these properties exactly, and
looks reasonably like a galaxy.  Solution not unique, an ensemble
generated.
SpaghettiLens is build on top of GLASS \citep{Lubini2012}, that builds and improves upon PixeLens \citep{Saha2004}.

\spl then tries to find a mass distribution $\kappa(x,y)$ that reproduces the input parameters.
Since there is no unique solution, \spl samples the solution space and produces an ensemble of solutions, as described in the paper \citep{Lubini2012}.

Step 3 by providing the visual feedback. Users can check the generated mass distribution and synthetic image.

It additinally provides plots of the modelled mass distribution and the contour lines of arrival time surface as feedback.

It provides direct visual feedback by rendering a synthetic image side
by side the original image, such that users can directly compare the
predictions of their model to the survey image. Explanation: Synthetic
image: A rendering of the derivative of the arrival time for each
pixel. This leads to a black and white image that reasonably looks
like the visual appearace of the model.

Using this direct visual feedback, even inexperienced users can
successfully model lenses using a iterative approach if they know what
to look for.

In the next step, several volunteers can work together / exchange their ideas and models and try to improve previous modelling attempts by others.
This leads to a set of models organised in a branching, tree like structre for each model.
Volunteers then organise them selves to narrow down the different branches and come up with a few consensus models in the end.

>>>>

\spl implements the ideas of the arrival-time surface into a modelling
program.

The first modelling step with \spl is for the user to call up the \sw
of interest, and make an educated guess for the topography of the
arrival-time surface --- specifically, the locations of the maxima,
minima and saddle points that correspond to images of the brightest
part of the source.  Input is given by tracing a proposal for the
saddle-point contours.  Figure \ref{fig:input-spag} shows some example
inputs --- and incidentally also indicates the origin of the name
\spl.

@@

In the second modelling step, once the user has sketched a proposed
spaghetti configurations, \spl sends the input to its server-side
modelling engine, called GLASS.  The task of GLASS is to find a mass
distribution $\kappa(x,y)$ that exactly reproduces the locations of
the maximum, minima and saddle points.  Now, this criterion by itself
is extremely underdetermined --- there are infinitely many mass
distributions that will reproduce a given set of maxima, minima and
saddle points, but typically they (a)~produce lots of extra images,
and (b)~look very unlike galaxies.  Additional assumptions (a prior)
are necessary.  GLASS uses a prior based on suggestions by by
\cite{1997MNRAS.292..148S}.
\begin{enumerate}
\item The mass distribution is built out of non-negative tiles of
  mass.  (Sometimes these tiles are called mass pixels, but we should
  emphasize that they are unrelated to image pixels, and are much
  larger.)
\item There is a notional lens centre, say $(x_0,y_0)$ which is
  identified with the maximum of the arrival time.  The source can
  have an arbitrary offset with respect to the lens centre.
\item The mass distribution must be centrally concentrated, in two
  respects.  First, the circularly averaged density must fall away
  like $$ \left[(x-x_0)^2+(y-y_0)^2\right]^{-1/2}$$ or more steeply.
  Second, the direction of increasing density at any $(x,y)$ can point
  at most $45^\circ$ away from from $(x_0,y_0)$.
\item The lens must symmetric with respect to $180^\circ$ rotations
  about $(x_0,y_0)$.  This symmetry assumption can be relaxed if the
  user wishes.
\end{enumerate}
There are still infinitely many models that satisfy both data and
prior constraints, but now they are more credible as galaxy lenses.
It is then possible to generate an ensemble of models.  The sampling
technique used by GLASS is described in \citep{Lubini2012}, and
improves upon earlier techniques \citep{2000AJ....119..439W,Saha2004}.
Typically, ensembles of 200 models are used.  That is to say, what we
call a \spl model is really the mean of an ensemble of 200 models, and
its estimated uncertainty is the range covered by the whole ensemble.

In the third modelling step, having generated a model ensemble, \spl
postprocesses it to present results and diagnostics to the user for
inspection.  This takes the form of three figures.
\begin{enumerate}
\item A grayscale plus contour map of the mass distribution.
\item A contour map of the arrival-time surface.
\item A synthetic image of the lensed features.
\end{enumerate}
After examining this feedback, the user can archive the results for
discussion, or modify their input and try again, or discard the
attempt altogether.

The fourth modelling step is discussion among modellers and iteration
on the model.  Any archived model can be revised by another user: they
can modify the spaghetti configuration slightly or drastically, or
change options like the size of the mass tiles.  Particularly
interesting lens candidates lead to trees of models in this way.
Forum discussion then prunes the tree, focussing attention on one or a
few models.
